Make any stupid plot you want

### 2021-01-19 draft/sections/math.tex
fleshed out fiber further, unpacked into set notation

\subsubsection{Variable Field in F }
\label{sec:data_fiber}
The fiber is a topological space containing the set of all observations that could possible be in the dataset, even values that do not actually exist in the dataset. For example, lets construct a fiber bundle that is just the temperature column from GHCN. Then the fiber $F$ is the range of physically possible temperature values $\text{Temp} \subset \mathbb{R}$. Lets rename this
\begin{equation}
F_0 = \{\text{Temp}\}
\end{equation}
so that more variables can be added to the fiber. For example, lets add the time from GHCN $F_1 =\text{Time}$
\begin{equation}
F =  \{\text{Temp}\} \times \text{Time}
\end{equation}

The elements in a given fiber set do not need to be scalers; for example the location can be encoded as $F_3=\text{Loc}$ where $(\text{lat},\text{lon{}) \in \text{Loc}$.  We add the location to the fiber by adding the set to the cartesian product
\begin{equation}
    F =  \{\text{Temp}\} \times \text{Time} \times \text{Loc}
\end{equation}

We also add the variable name and measurement scale $M$ monoid to the fiber $F_i$ such that 
\begin{align}
    F_0 &= (\mathrm{temperature}, \text{Temp}, M_0)\\
    F_1 &= (\mathrm{time}, \text{Time}, M_1)\\
    F_2 &= (\mathrm{location}, \text{Loc}, M_2)
\end{align}
 or more generally (name, set, monoid)
 \begin{equation}
    F_i = (\mathrm{name}, \{x_{i}, \ldots, \x_{n}\}, M_{i})
 \end{equation}
as described in Spivak's categorical formulation of databases \cite{spivakSIMPLICIALDATABASES}. We describe $M$ in more detail in section~\ref{sec:data_monoid}. We can keep growing the fiber
\begin{equation}
F= F_{0} \times \ldots \times F_{n}
\end{equation}
while also retaining the individual sets $F_{i}$ that we will transform into visual characteristics of the graphic.

@@

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

@@

\begin{figure}[h!]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}
The topology $K$ and the

@@

\subsubsection{Example: Temperature}
\textcolor{red}{Moved \& walked through b/c was getting clunky to not have terms yet}\\

@@

\begin{figure}[h!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
        %% add box around neighboring P and Map
        \label{fig:base_example_line}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2k.png}
        \label{fig:base_example_plane}
    \end{subfigure}
    \label{fig:base_example}
    \caption{These two datasets have the same fiber of temperature but different base spaces. In figure~\ref{fig:base_example_line} the temperature values are 1D continuous, while in figure~\ref{fig:base_example_plane} the temperature values are 2D continuous.}
\end{figure}

@@

\begin{figure}[h!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2f.png}
        \label{fig:fiber_example_plane}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_3f.png}
        \label{fig:fiber_example_cube}
    \end{subfigure}
    \label{fig:fiber_example}
    \caption{The fiber is expanded to include metadata fields that describe the semantics of $K$. In figure~\ref{fig:fiber_example_plane} the fiber is \textrm{temperature} $\times$ \textrm{time} and in figure~\ref{fig:fiber_example_cube} the fiber is \textrm{temperature}$\times$ \textrm{latitude} $\times$ {longitude}}.
\end{figure}

@@

\begin{figure}[h!]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \label{fig:fiber_example_section}
    \caption{The section $\tau_1$ returns the blue points,  while $\tau_2$ returns the purple points. $\Gamma(E)$ is the set of all sections, including $\tau_1$ and $\tau_2$}.  
\end{figure}
Given the fiber described in equation~\ref{eq:fiber_plane}, the sections $\tau_{1}$ and $\tau_{2}$ in figure~\ref{fig:fiber_example_section} return tuples of the form
\begin{equation}
\tau(k) = (k, (temperature, time))
\end{equation}
such that sections with the constraint that time is monotonic return a timeseries.

@@

\subsubsection{{Example}}
\begin{figure}[h!]
    \includegraphics[width=.4\linewidth]{figures/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

@@

\begin{equation}
    \label{eq:artist}
    \begin{tikzcd}
        E \arrow[r, "\nu"] \arrow[rd, "\pi"'] & V \arrow[d, "\pi"] & \xi^*V \arrow[r, "Q"] \arrow[d, "\xi^*\pi"'] \arrow[l, "\xi^*"'] & H \arrow[ld, "\pi"] \\
                                              & K                  & S \arrow[l, "\xi"']                                              &                    
        \end{tikzcd}
\end{equation}
The pullback map $\xi^*$ copies each value in $V$ over $k$ to $s$ in corresponding $S_k$ such that $\xi^*V$ can have multiple values that map to one value in $V$.

@@

\subsubsection{Assembling Marks}
\label{sec:artist_q}
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}  
\end{figure}

@@

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \label{fig:artist_mark_change}
\end{figure}

@@

\paragraph{Q: scatter plot}
\\
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/scatter.png}
    \label{fig:artist_scatter}
\end{figure}

@@

\paragraph{Q: line plot}
\\
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \label{fig:artist_line}
\end{figure}

@@

\paragraph{Q: heatmap}
\\
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/heatmap.png}
    \label{fig:artist_heatmap}
\end{figure}
The heatmap in figure~\ref{fig:artist_heatmap} 
\begin{equation}
Q(xpos, ypos, color)
\end{equation}
has in the simple case a direct lookup into $K$ to obtain the $\mu = (x,y,c)$ values that are mapped into 
(arrays are upside down, x is y, is x) - $\xi$ is about translating indices from data to visual alignment
\begin{align}
D_{RGB} &= color(\xi(\alpha_1, \alpha_2))
D_x & = xpos(\xi(\alpha_1, \alpha_2))
D_y &= ypos(\xi(\alpha_1, \alpha_2))
\end{align}
through $\rho$.

@@

\begin{figure}[h!]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) observations to all observations are connected to at least 3 other observations}
    \label{fig:triangle_simplex}
\end{figure}

@@

\subparagraph{Example}
\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \label{fig:triangle_torus}
\end{figure}

>>>>

\subsubsection{Variable Field in F }
\label{sec:data_fiber}
The fiber is a topological space containing the set of all observations that could possible be in the dataset, even values that do not actually exist in the dataset. For example, lets construct a fiber bundle that is just the temperature column from GHCN. Then the fiber $F$ is the range of physically possible temperature values $\{tp_0, \ldots, tp_n\} \subset \mathbb{R}$. Lets rename this
\begin{equation}
F_0 = \{tp_0, \ldots, tp_n\} 
\end{equation}
so that more variables can be added to the fiber. Then lets add the time from GHCN $F_1 =\{t_0, \ldots, t_n\}$
\begin{equation}
F =  \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} 
\end{equation}

The elements in a given fiber set do not need to be scalers; for example the location can be encoded as the set of tuples $(\textit{lat},\textit{lon})$.  We add the location to the fiber 
\begin{equation}
    F = \{tp_0, \ldots, tp_n\}  \times \{ti_0, \ldots, ti_n\} \times 
    \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}
\end{equation}
such that one record $r$ in the fiber would have the form $(tp, t, (la,ln))$. We also add the variable name and measurement scale $M$ monoid to the fiber $F_i$ such that 
\begin{align}
    F_0 &= (\textit{temperature}, \{tp_0, \ldots, tp_n\}, M_0)\\
    F_1 &= (\textit{time}, \{ti_0, \ldots, ti_n\}, M_1)\\
    F_2 &= (\textit{location},   \{(la_0, ln_0),\ldots, (la_j, ln_k), \ldots (la_n,ln_n)\}, M_2)
\end{align}
 or more generally (name, set, monoid)
 \begin{equation}
    F_i = (\textit{name}, \{x_{i}, \ldots, x_{n}\}, M_{i})
 \end{equation}
as described in Spivak's categorical formulation of databases \cite{spivakSIMPLICIALDATABASES}. We describe $M$ in more detail in section~\ref{sec:data_monoid}. We can keep growing the fiber
\begin{equation}
F= F_{0} \times \ldots \times F_{n}
\end{equation}
while also retaining the individual sets $F_{i}$ that we will transform into visual characteristics of the graphic.

@@

\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/k_different_types.png}
    \label{fig:base_space_types}
    \caption{The topological base space $K$ encodes the connectivity of the data space, for example if the data is independent points or a map or on a sphere}
\end{figure}

@@

\begin{figure}[ht!]
    \label{fig:base_space_div}
    \includegraphics[width=.5\linewidth]{figures/math/k_qspace.png}
    %%include second figure where fiber goes otherway
\end{figure}
The topology $K$ and the

@@

\subsubsection{Case Study: Temperature}
\textcolor{red}{Moved \& walked through b/c was getting clunky to not have terms yet}\\

@@

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_1k.png}
        %% add box around neighboring P and Map
        \label{fig:base_example_line}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2k.png}
        \label{fig:base_example_plane}
    \end{subfigure}
    \label{fig:base_example}
    \caption{These two datasets have the same fiber of temperature but different base spaces. In figure~\ref{fig:base_example_line} the temperature values are 1D continuous, while in figure~\ref{fig:base_example_plane} the temperature values are 2D continuous.}
\end{figure}

@@

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_2f.png}
        \label{fig:fiber_example_plane}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/temp_3f.png}
        \label{fig:fiber_example_cube}
    \end{subfigure}
    \label{fig:fiber_example}
    \caption{The fiber is expanded to include metadata fields that describe the semantics of $K$. In figure~\ref{fig:fiber_example_plane} the fiber is \textrm{temperature} $\times$ \textrm{time} and in figure~\ref{fig:fiber_example_cube} the fiber is \textrm{temperature}$\times$ \textrm{latitude} $\times$ {longitude}}.
\end{figure}

@@

\begin{figure}[ht!]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \label{fig:fiber_example_section}
    \caption{The section $\tau_1$ returns the blue points,  while $\tau_2$ returns the purple points. $\Gamma(E)$ is the set of all sections, including $\tau_1$ and $\tau_2$}.  
\end{figure}
Given the fiber described in equation~\ref{eq:fiber_plane}, the sections $\tau_{1}$ and $\tau_{2}$ in figure~\ref{fig:fiber_example_section} return tuples of the form
\begin{equation}
\tau(k) = (k, (temperature, time))
\end{equation}
such that sections with the constraint that time is monotonic return a timeseries.

@@

\subsubsection{{Example}}
\begin{figure}[ht!]
    \includegraphics[width=.4\linewidth]{figures/math/render.png}
    \caption{}
    \label{fig:render}
\end{figure}

@@

\begin{equation}
    \label{eq:artist_daigram}
    \begin{tikzcd}
        E \arrow[r, "\nu"] \arrow[rd, "\pi"'] & V \arrow[d, "\pi"] & \xi^*V \arrow[r, "Q"] \arrow[d, "\xi^*\pi"'] \arrow[l, "\xi^*"'] & H \arrow[ld, "\pi"] \\
                                              & K                  & S \arrow[l, "\xi"']                                              &                    
        \end{tikzcd}
\end{equation}
The pullback map $\xi^*$ copies each value in $V$ over $k$ to $s$ in corresponding $S_k$ such that $\xi^*V$ can have multiple values that map to one value in $V$.

@@

\subsubsection{Assembling Marks}
\label{sec:artist_q}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}  
\end{figure}

@@

\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/diff_type_q.png}
    \label{fig:artist_mark_change}
\end{figure}

@@

\paragraph{Q: scatter plot}\mbox{} \\
\label{sec:artist_example_scatter}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/scatter.png}
    \label{fig:artist_scatter}
\end{figure}

@@

\paragraph{Q: line plot}\mbox{} \\
\label{sec:artist_example_line}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/line.png}
    \label{fig:artist_line}
\end{figure}

@@

\paragraph{Q: heatmap}\mbox{} \\
\label{sec:artist_example_heatmap}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/heatmap.png}
    \label{fig:artist_heatmap}
\end{figure}
The heatmap in figure~\ref{fig:artist_heatmap} 
\begin{equation}
Q(xpos, ypos, color)
\end{equation}
has in the simple case a direct lookup into $K$ to obtain the $\mu = (x,y,c)$ values that are mapped into 
(arrays are upside down, x is y, is x) - $\xi$ is about translating indices from data to visual alignment
\begin{align}
D_{RGB} &= color(\xi(\alpha_1, \alpha_2))
D_x & = xpos(\xi(\alpha_1, \alpha_2))
D_y &= ypos(\xi(\alpha_1, \alpha_2))
\end{align}
through $\rho$.

@@

\begin{figure}[ht!]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) observations to all observations are connected to at least 3 other observations}
    \label{fig:triangle_simplex}
\end{figure}

@@

\subparagraph{Example}
\begin{figure}[ht!]
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \label{fig:triangle_torus}
\end{figure}
